\section{\texorpdfstring{$\omega$}{omega}-Hausdorff rank}

In this chapter, we focus on the special case of the Hausdorff rank associated with the ordinal $\omega$. This case is of particular interest due to its connections with countable structures and its role in the classification of infinite linear orders. We introduce new families of properties and analyze their relationships, providing tools that will be essential for the study of types and decidability.

\begin{definitions}\label{bounded-definitions}
  Let $\alpha > 0$ be an ordinal.

  We define:
  \begin{enumerate}
    \item $\mathcal{S}^{1}_{\alpha} := \qqo^{< \alpha}$
    \item $\mathcal{S}^{\omega}_{\alpha} := \rb{\qqo^{< \alpha}} \setminus \lb{\qqo^{< \alpha}}$
    \item $\mathcal{S}^{\agemo}_{\alpha} := \lb{\qqo^{< \alpha}} \setminus \rb{\qqo^{< \alpha}}$
    \item $\mathcal{S}^{\oo}_{\alpha} := \bounded{\qqo^{< \alpha}} \setminus \ps { \lb {\qqo^{< \alpha}} \cup \rb{\qqo^{< \alpha}} }$
  \end{enumerate}

  The names will soon be justified.
\end{definitions}

\begin{lemma}\label{bounded-structure}
  Let $\alpha > 0$ be an ordinal.

  Then,
  \begin{enumerate}
    \item $\rb{\qqo^{< \alpha}} = \sum_{\omega}{\qqo^{< \alpha}}$.
    \item $\lb{\qqo^{< \alpha}} = \sum_{\agemo}{\qqo^{< \alpha}}$.
    \item $\bounded{\qqo^{< \alpha}} = \sum_{\oo}{\qqo^{< \alpha}}$.
  \end{enumerate}
\end{lemma}

\begin{proof}
  1. Let us prove the first part.
  ($\supseteq$) Let $M \in \sum_{\omega}{\qqo^{< \alpha}}$ be a linear order.

  Let $M = \sum_{i \in \omega} M_i$ be the decomposition of $M$,
  where $M_i \in \qqo^{< \alpha}$.

  Let $x, y \in M$ be any two points in $M$. WLOG $x \le y$.

  Suppose $x \in M_i$ and $y \in M_j$ for $i, j \in \omega$.

  Since $i$ and $j$ have a finite distance in $\omega$,
  we conclude $[x, y] \subseteq M_i + \ldots + M_j$,
  and thus $[x, y] \subseteq \ps{\qqo^{< \alpha}}^+ = \qqo^{{\alpha}}$.

  ($\subseteq$) Let $M \in \rb{\qqo^{< \alpha}}$ be a linear order.

  Since $M$ is countable, let $\set{x_i}_{i \in \omega} M$ be a right cofinal
  $\omega$-sequence in $M$.

  Let $M_0 = (-\infty, x_0]$ and and $M_i = (x_{i-1}, x_{i}]$ for $i > 0$.

  Then $M = \sum_{i \in \omega} M_i$.

  But $M_i$ is a right-bounded interval and thus $M_i \in \qqo^{< \alpha}$, so $M \in \sum_{\omega}{\qqo^{< \alpha}}$.

  2. The second part is symmetric.

  3. The third part follows from~\cref{bounded-is-right-plus-left}:
  \begin{align*}
    \bounded{\qqo^{< \alpha}}
     & = \lb{\qqo^{< \alpha}} + \rb{\qqo^{< \alpha}}                     \\
     & = \sum_{\agemo}{\qqo^{< \alpha}} + \sum_{\omega}{\qqo^{< \alpha}} \\
     & = \sum_{\oo}{\qqo^{< \alpha}}
  \end{align*}
\end{proof}

% Discussion
The next lemma describes the structure of right-bounded, left-bounded, and bounded intervals in terms of sums over $\omega$
and related index sets. This structural insight is crucial for later results on types and expressibility.

An immediate corollary of~\cref{bounded-structure} is that:
\begin{corollary}\label{le-alpha-corollary}
  Let $\alpha > 0$ be an ordinal.

  Then,
  \begin{enumerate}
    \item $\qqo^{\le \alpha} = \ps{\bounded{\qqo^{< \alpha}}}^+$
    \item $\qqo^{= \alpha}
            = \ps{
              \mathcal{S}^{\omega^\ast}_{\alpha}
              \uplus \mathcal{S}^{\omega}_{\alpha}
            }^+$
  \end{enumerate}
\end{corollary}

\begin{lemma}\label{shape-structure-successor}
  Let $\alpha$ be an ordinal.

  Let $s \in \set{\omega, \agemo, \oo}$.

  Then, we have the following:

  \[
    \mathcal{S}^{s}_{\alpha + 1} = \sum_{i \in s}{\qqo^{= \alpha}}
  \]
\end{lemma}

\begin{proof}
  It is enough to prove the case $s = \omega$,
  since $s = \agemo$ follows by symmetry,
  and $s = \oo$ follows by adding the previous two cases.

  ($\subseteq$) Let $M \in \mathcal{S}^{\omega}_{\alpha + 1}$.

  By~\cref{bounded-structure} and~\cref{le-alpha-corollary}
  \[
    \rb{\qqo^{< \alpha + 1}}
    = \sum_{\omega}{\qqo^{< \alpha + 1}}
    = \sum_{\omega}{\qqo^{\le \alpha}}
    = \sum_{\omega}{\ps{\bounded{\qqo^{< \alpha}}}^+}
    = \sum_{\omega}{\bounded{\qqo^{< \alpha}}}
  \]

  since by~\cref{bounded-definitions}, $M \in \rb{\qqo^{< \alpha + 1}}$,
  we conclude that $M = \sum_{i \in \omega}{M_i}$ for a sequence
  $\set{M_i}_{i \in \omega} \subseteq \bounded{\qqo^{< \alpha}}$.

  If $M_i \in \qqo^{= \alpha}$ held for only finitely many $i \in \omega$,
  we would have $M \in \qqo^{\le \alpha}$, which is a contradiction
  to $M \notin \lb{\qqo^{< \alpha}}$.

  Thus, $M_i \in \qqo^{= \alpha}$ holds
  for infinitely many $i \in \omega$, and thus (by adjoining $M_i \in \qqo^{< \alpha}$
  to the next $\qqo^{= \alpha}$ one) we conclude $M \in \sum_{\omega}{\qqo^{= \alpha}}$.

  ($\supseteq$) Let $M \in \sum_{\omega}{\qqo^{= \alpha}}$.

  Since $M \in \sum_{\omega} {\qqo^{< \alpha + 1}}$, by~\cref{bounded-structure},
  $M \in \rb{\qqo^{< \alpha + 1}}$.

  By~\cref{le-alpha-corollary},
  \[
    M \in \sum_{\omega}{\qqo^{= \alpha}}
    = \sum_{\omega} \ps{ \mathcal{S}^{\omega^\ast}_{\alpha} \uplus \mathcal{S}^{\omega}_{\alpha} }^+
    = \sum_{\omega} \ps{ \mathcal{S}^{\omega^\ast}_{\alpha} \uplus \mathcal{S}^{\omega}_{\alpha} }
  \]

  Suppose $M = \sum_{i \in \omega}{M_i}$ where $M_i \in \mathcal{S}^{s_i}_{\alpha}$
  for $s_i \in {\omega^\ast, \omega}$. By the pigeonhole principle,
  suppose eventually $s_i = s$. WLOG, eventually $s_i = \omega$. WLOG,
  $s_i = \omega$ for all $i \in \omega$, so $M \in \sum_{\omega} \mathcal{S}^{\omega}_{\alpha}$.

  Suppose by contradiction $M = \qqo^{\le \alpha} = \ps{\bounded{\qqo^{< \alpha}}}^+$.
  In particular, by the pigeonhole principle, there exists some $N \in \omega$
  such that $\sum_{N \le i < \omega} M_i \in \bounded{\qqo^{< \alpha}}$,
  which is a contradiction because it follows that $M_{N+1} \in \qqo{< \alpha}$ as it is bounded
  between $M_N$ and $M_{N+2}$.
\end{proof}

\begin{lemma}\label{shape-structure-limit}
  Let $\set{\alpha_i}_{i \in \omega}$ be a strictly
  increasing ordinal sequence, and let $\alpha = \sup_{\alpha_i}$.

  Then,
  \[
    \mathcal{S}^{s}_{\alpha} = \sum_{i \in s}{\qqo^{[\alpha_i, \alpha)}}
  \]
\end{lemma}

\begin{proof}
  Again, it is enough to prove for $s = \omega$.
  ($\subseteq$) Let $M \in \mathcal{S}^{\omega}_{\alpha}$.
  Let ${y_i}_{i < \omega}$ be a right cofinal $\omega$-sequence in $M$.

  Thus we can choose some $x_0$ far enough such that $(-\infty, x_0] \in \qqo^{[\alpha_0, \alpha)}$,
  and $x_0 > y_0$.
  Now by induction we choose $x_1$ such that $(x_0, x_1] \in \qqo^{[\alpha_1, \alpha)}$,
  and $x_1 > y_1$.

  By iterating $\omega$ times we get an $\omega$-sequence $\set{M_i}_{i \in \omega}$
  such that $M = \sum_{i \in \omega} M_i$ and $M_i \in \qqo^{[\alpha_i, \alpha)}$,
  where $M_i = (x_{i-1}, x_i]$ (where $x_{-1} := -\infty$).


  ($\supseteq$) Let $M \in \sum_{i \in \omega}{\qqo^{[\alpha_i, \alpha)}}$.
  It is obvious that $M \in \rb{\qqo^{< \alpha}}$ since every right-bounded
  ray is in $\qqo^{\le \alpha_i}$ for some $i \in \omega$.

  However, $M \notin \qqo^{< \alpha_i}$ for any $i \in \omega$, so $M \notin \qqo^{< \alpha}$.


\end{proof}

\begin{lemma}\label{shape-structure}
  Let $\set{\alpha_i}_{i \in \omega}$ be a non-decreasing ordinal sequence,
  and let $\alpha = \sup_{\alpha_i + 1}$.

  Then,
  \[
    \mathcal{S}^{s}_{\alpha} = \sum_{i \in s}{\qqo^{= \alpha_i}}
  \]
\end{lemma}

\begin{proof}
  It is just a way to write~\cref{shape-structure-successor} and~\cref{shape-structure-limit}
  together more succinctly.
\end{proof}

\begin{note}
  For the proof of~\cref{shape-structure}, we actually use the fact that
  we work over $\qqo = \Gamma_{\omega}$. This proof would not have worked over $\Gamma_{\beta}$
  for $\beta > \omega$.
\end{note}

% Discussion
The corollary that follows gives an immediate application of the previous lemma, relating the $\le \alpha$ property to
bounded sums.

\begin{corollary}
  Let $\alpha, \delta > 0$ be limit ordinals.

  Let $s \in \set{\omega, \agemo, \oo}$

  Then,

  \[
    \mathcal{S}^{s}_{\alpha + \delta}
    = \sum_{\mathcal{S}^{s}_{\delta}}{\qqo^{= \alpha}}
  \]
\end{corollary}

\begin{proof}
  Suppose that $\delta = \sup_{i \in s} \delta_i$,
  where $\set{\delta_i}_{i \in s}$ is an increasing $s$-sequence
  of ordinals.

  Then $\alpha + \delta = \sup_{i \in s} \ps{\alpha_i + \delta_i}$.

  \[
    \mathcal{S}^{s}_{\alpha + \delta}
    = \sum_{i \in s}{\mathcal{S}^{s}_{\alpha + \delta_i}}
    = \sum_{i \in s}\sum_{\qqo^{= \delta_i}}{\qqo^{= \alpha}}
    = \sum_{\sum_{i \in s}{\qqo^{= \delta_i}}}{\qqo^{= \alpha}}
    = \sum_{\mathcal{S}^{s}_{\delta}}{\qqo^{= \alpha}}
  \]
\end{proof}
